Type Evolution in Programming Systems

### 2024-05-29 paper.tex
Data wrangling references

\paragraph{Live programming.}
Live programming~\cite{tanimoto90, rein2018exploratory} seeks to speed up feedback in (mostly solo) programming.
% \cite{mcdirmid13, Rauch_2019, beckmann2021shortening}
% A special case of live programming is \emph{hot reloading} or \emph{hot swapping})~\cite{barenz2020essence, hicks2005dynamic} that installs code changes into a running system preserving state but not attempting to migrate that state through type changes.
\citet{SemanticDeltas} proposes a research agenda for thoroughly live DSL programming focusing on \emph{Semantic Deltas} that unify edit-time and run-time change.
\citet{RuntimeConstraint} study the live modeling problem in Section \ref{livemodeling}. They use constraints supplied by the developer to solve for the best new runtime state following a DSL language evolution. \citet{vanRozen19} take an alternative approach to the same problem using \emph{origin tracking} in textual editing.
%\citet{vanRozen23} subsequently proposed a metaprogramming language for live DSL programming.

>>>>

\paragraph{Data wrangling.} The evolution of data formats is a concern in Data Science. Data~Diff~\cite{sutton18} compares two versions of tabular data and finds a minimal matching transformation.
Wrangler~\cite{kandel11} observes interactive manipulation of sampled data to write data transformation programs.
\citet{petersohn20} present an algebra of operations on Python dataframes that include shape transformations.

\paragraph{Live programming.}
Live programming~\cite{tanimoto90, rein2018exploratory} seeks to speed up feedback in (primarily solo) programming.
% \cite{mcdirmid13, Rauch_2019, beckmann2021shortening}
\emph{Hot reloading/swapping}~\cite{barenz2020essence, hicks2005dynamic} installs code changes into a running system preserving some runtime state but not attempting to migrate that state through type changes.
\citet{SemanticDeltas} proposes a research agenda for thoroughly live DSL programming focusing on \emph{Semantic Deltas} that unify edit-time and run-time change.
\citet{RuntimeConstraint} study the live modeling problem in Section \ref{livemodeling}. They use constraints supplied by the developer to solve for the best new runtime state following a DSL language evolution. \citet{vanRozen19} take an alternative approach to the same problem using \emph{origin tracking} in textual editing.
%\citet{vanRozen23} subsequently proposed a metaprogramming language for live DSL programming.
